Spatio-temporal estimates of HIV risk group proportions for adolescent girls and young women across 13 priority countries in sub-Saharan Africa

### 2022-09-28 src/docs_28-09-22-toronto/28-09-22-toronto.Rmd
Iterate Toronto talk

* Consider models of the form
\begin{align}
    y_{itak}^\star &\sim \text{xPoisson}(\lambda_{itak}) \\
    \log(\lambda_{itak}) &= \theta_{ita} + \beta_k + \zeta_{c[i]k} + \alpha_{ac[i]k} + \phi_{ik} + \gamma_{tk} + \delta_{itk}.
\end{align}
* With the following random effects
  * $\theta_{ita}$: observation (IID)
  * $\beta_k$: category (IID)
  * $\zeta_{ck}$: country-category (IID x IID) 
  * $\alpha_{ack}$: age-country-category (IID x IID x IID)
  * $\phi_{ik}$: space-category ({IID, **Besag**} x IID)
  * $\gamma_{tk}$: year-category ({**IID**, AR1} x IID)
  * $\delta_{itk}$: space-year-category (Implemented but crashing on cluster at the moment...)

@@

* Consider models of the form
\begin{align}
    y_{ia4}^\star &\sim \text{Binomial} \left( y_{ia3}^\star + y_{ia4}^\star, q_{ia} \right), \label{eq:logistic-regression} \\
    q_{ia} &= \text{logit}^{-1} \left( \eta_{ia} \right), \\
    \eta_{ia} &= \beta_0 +  \zeta_{c[i]} + \alpha_{ac[i]} + \phi_i + \beta_{\texttt{cfsw}} \texttt{cfsw}_{c[i]}.
\end{align}
* With the following terms:
  * $\beta_0$: intercept
  * $\zeta_{c[i]}$: country effects (IID)
  * $\alpha_{ac[i]}$ age-country effects (IID)
  * $\phi_i$: spatial effects (IID, **Besag**)
  * Clients of FSW covariates ($\texttt{cfswever}$, $\texttt{cfswrecent}$) [@hodgins2022population]

@@

* Take 1000 samples from each model, then multiply suitably to generate estimates for all four risk groups
* We adjusted the $k = 4$ risk group to match age-country FSW estimates
  * Obtained these by disaggregating @stevens2022estimating by age using estimates of sexually active population from @nguyen2022trends

@@

@@

```{r echo=FALSE, out.width="350px", fig.cap = "Here is another view of the discontinuity."}
knitr::include_graphics("depends/aids-abstract-B.pdf")
```

@@

* Integration of all relevant surveys
  <!-- * Modular approach allowed estimating FSW proportion even for surveys without a specific transactional sex question -->
* Alleivating small-sample sizes by borrowing information
  <!-- * We borrowed information across space, between countries and over surveys so that our estimates more plausibly reflect spatial heterogeneity -->
* Estimates where there isn't direct data
  <!-- * Although some people think of this as "making up data", the data almost never "speak for themselves" (everything is a model) -->
  <!-- * Uncertainty should be higher in regions with infilling: important to transparently communicate this -->

@@

## HIV incidence by risk group

* Risk group proportion estimates plus relative risk ratio estimates to disaggregate general population HIV incidence estimates
\begin{align*}
I_{ia} &= \sum_k \lambda_{iak} N_{iak} \\
&= \sum_k \lambda_{ia2} \text{RR}_k N_{iak}.
\end{align*}

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of HIV incidence $\lambda_{iak}$ and number of new HIV infections $I_{iak}$ by district, age group and risk group
\end{tcolorbox}
\end{center}

@@

```{r echo=FALSE, out.width = "300px", fig.cap = "Mozambique stands out."}
knitr::include_graphics("depends/cohabiting-girls.pdf")
```

@@

* Simplistic infections reached analysis
  <!-- * No accounting for difficulty in reaching each strata -->
  <!-- * Variable intervention effectiveness -->
  <!-- * Change in strata membership -->
* Under-reporting of high risk sexual behaviours
  <!-- * Variation in under-reporting (likely by age, foremost, and location, less so) particularly concerning -->
* Risk groups definition justification not clear
  <!-- * Didn't consider other important characteristics that may determine risk e.g. condom usage -->
* Only focused on AGYW 15-29
  <!-- * Could be extended to adults of both sexes aged 15-49 -->

@@

* Risk group estimates can help implement the Global AIDS Strategy; tool and user guide currently being prepared!
* Importance of reaching FSW
* Countries have different epidemic profiles

>>>>

* Consider models of the form
\begin{align*}
    y_{itak}^\star &\sim \text{xPoisson}(\lambda_{itak}) \\
    \log(\lambda_{itak}) &= \theta_{ita} + \beta_k + \zeta_{c[i]k} + \alpha_{ac[i]k} + \phi_{ik} + \gamma_{tk} + \delta_{itk}.
\end{align*}
* The terms are
  * $\theta_{ita}$: observation (IID)
  * $\beta_k$: category (IID)
  * $\zeta_{ck}$: country-category (IID x IID) 
  * $\alpha_{ack}$: age-country-category (IID x IID x IID)
  * $\phi_{ik}$: space-category ({IID, **Besag**} x IID)
  * $\gamma_{tk}$: year-category ({**IID**, AR1} x IID)
  * $\delta_{itk}$: space-year-category (Implemented but crashing on cluster at the moment...)

@@

* Consider models of the form
\begin{align*}
    y_{ia4}^\star &\sim \text{Binomial} \left( y_{ia3}^\star + y_{ia4}^\star, q_{ia} \right), \label{eq:logistic-regression} \\
    q_{ia} &= \text{logit}^{-1} \left( \eta_{ia} \right), \\
    \eta_{ia} &= \beta_0 +  \zeta_{c[i]} + \alpha_{ac[i]} + \phi_i + \beta_{\texttt{cfsw}} \texttt{cfsw}_{c[i]}.
\end{align*}
* The terms are
  * $\beta_0$: intercept
  * $\zeta_{c[i]}$: country effects (IID)
  * $\alpha_{ac[i]}$ age-country effects (IID)
  * $\phi_i$: spatial effects (IID, **Besag**)
  * Clients of FSW covariates (`cfswever`, **`cfswrecent`**) [@hodgins2022population]

@@

* Take 1000 samples from each model, then manipulate suitably to generate estimates for all four risk groups
* We adjusted the samples from the $k = 4$ risk group to match age-country FSW estimates, reallocating into non-regular partner(s)
  * Obtained these by disaggregating @stevens2022estimating by age using estimates of sexually active population from @nguyen2022trends

@@

```{r echo=FALSE, out.width="350px", fig.cap = "Results of FSW age disaggregation."}
knitr::include_graphics("depends/age-disagg-fsw-line.pdf")
```

```{r echo=FALSE, out.width="350px", fig.cap = "PIT histograms and ECDF difference plots."}
knitr::include_graphics("depends/coverage.pdf")
```

@@

```{r echo=FALSE, out.width="350px", fig.cap = "Viewing the discontinuity another way."}
knitr::include_graphics("depends/aids-abstract-B.pdf")
```

```{r echo=FALSE, out.width="350px", fig.cap = "Proportions of variance explained."}
knitr::include_graphics("depends/aaa-variance-proportions.pdf")
```

@@

* Integration of all relevant surveys
  * Two-stage approach allowed estimating FSW proportion even for surveys without a specific transactional sex question
* Alleivating small-sample sizes by borrowing information
  * We borrowed information across space, between countries and over surveys so that our estimates more plausibly reflect spatial heterogeneity
* Estimates where there isn't direct data
  * Although some people think of this as "making up data", the data almost never "speak for themselves" (everything is a model)
  * Uncertainty should be higher in regions with infilling: important to transparently communicate this

@@

## HIV prevalence and incidence by risk group

* We used our risk group proportion estimates together with incidence relative risk ratios and prevalence ratios to disaggregate general population HIV estimates in the most recent year
* Disaggregated number of new infections on a linear scale, and people living with HIV (PLHIV) on a logit scale
  * Using a linear scale for PLHIV resulted in prevalences outside $[0, 1]$

\begin{center}
\begin{tcolorbox}[width=0.9\textwidth, colframe={title}, colback={white}, title={}]
$\implies$ Estimates of HIV incidence $\lambda_{iak}$, number of new HIV infections $I_{iak}$, HIV prevalence $\rho_{iak}$ and PLHIV $H_{iak}$ by district, age group and risk group
\end{tcolorbox}
\end{center}

@@

@@

* Simplistic infections reached analysis
  * No accounting for difficulty in reaching each strata
  * Variable intervention effectiveness
  * Change in strata membership
* Under-reporting of high risk sexual behaviours
  * Variation in under-reporting (likely by age, foremost, and location, less so) particularly concerning
* Risk groups definition justification not clear
  * Didn't consider other important characteristics that may determine risk e.g. condom usage
* Only focused on AGYW 15-29
  * Could be extended to adults of both sexes aged 15-49

@@

* Risk group estimates can help implement the Global AIDS Strategy; tool and user guide currently being rolled out!
* Importance of reaching FSW
* Countries have different epidemic profiles
